### 2017-02-02 doc/manuscript.Rmd
Moved some tables and figs into Supplemental Results

tab_qic <- tab('qic', 'Comparison of GEE model fitness for variable selection using quasi-likelihood information criteria.')
tab_basic <- tab('basic', 'Basic characteristics of PROMISE participants at each of the 3 clinic visits.')
tab_tagfa <- tab('tagfa', 'Concentration (nmol/mL) and relative percent (mol%) values of triacylglycerol fatty acids in PROMISE participants at the baseline visit (2004-2006).')
tab_gee_unadj <- tab('gee_unadj', 'Raw estimates and confidence interval values for *time*-adjusted GEE models of the association of the triacylglycerol fatty acids (mol% and nmol/mL) and total clinically-measured TAG with insulin sensitivity and beta-cell function outcomes using the 6 year longitudinal data from the PROMISE cohort. Estimates represent a percent difference in the outcome per SD increase in the fatty acid. P-values were adjusted for the BH false discovery rate, with an asterisk (*) denoting a significant (p<0.05) association.')
tab_gee_adj <- tab('gee_adj', 'Raw estimates and confidence interval values for *fully*-adjusted GEE models of the association of the triacylglycerol fatty acids (mol% and nmol/mL) and total clinically-measured TAG with insulin sensitivity and beta-cell function outcomes using the 6 year longitudinal data from the PROMISE cohort. Variables controlled for were follow-up time, waist circumference, baseline age, ethnicity, sex, ALT, physical activity, and total NEFA. Estimates represent a percent difference in the outcome per SD increase in the fatty acid. P-values were adjusted for the BH false discovery rate, with an asterisk (*) denoting a significant (p<0.05) association.')

fig_consort <- fig('consort', 'CONSORT diagram of PROMISE participants over the 3 visits.')
fig_dag_is <- fig('dag_is', 'Directed acyclic graphic output from the DAGitty online software for insulin sensitivity.')
fig_dag_bcf <- fig('dag_bcf', 'Directed acyclic graphic output from the DAGitty online software for beta-cell function.')
fig_tagfa <- fig('tagfa', 'Distribution of the composition of triacylglycerol fatty acids in the baseline visit of PROMISE participants (2004-2006). Boxplots represent the median and interquartile range of the fatty acid values.')
fig_heatmap <- fig('heatmap', 'Pearson correlation heatmap of triacylglycerol fatty acids (nmol/mL) with continuous basic and metabolic characteristics of PROMISE participants from the baseline visit (2004-2006). Darker orange represents a positive correlation; darker blue represents a negative correlation.')
fig_heatmap_mol <- fig('heatmap_mol', 'Pearson correlation heatmap of triacylglycerol fatty acids (mol%) with continuous basic and metabolic characteristics of PROMISE participants from the baseline visit (2004-2006). Darker orange represents a positive correlation; darker blue represents a negative correlation.')
fig_heatmap_tagfa <- fig('heatmap_tagfa', 'Pearson correlation heatmap of the triacylglycerol fatty acids in the PROMISE participants from the baseline visit (2004-2006). The correlations of fatty acids grouped using heirarchical cluster analysis; fatty acids along the x and y axis are ordered according to this analysis. Darker orange represents a positive correlation; darker blue represents a negative correlation.')
fig_gee_unadj <- fig('gee_unadj', 'Time-adjusted GEE models of the association of the triacylglycerol fatty acids (mol% and nmol/mL) and total clinically-measured TAG with insulin sensitivity and beta-cell function outcomes using the 6 year longitudinal data from the PROMISE cohort. X-axis values represent a percent difference in the outcome per SD increase in the fatty acid. P-values were adjusted for the BH false discovery rate, with the largest dot representing a significant (p<0.05) association.')
fig_gee_adj <- fig('gee_adj', 'Fully-adjusted GEE models of the association of the triacylglycerol fatty acids (mol% and nmol/mL) and total clinically-measured TAG with insulin sensitivity and beta-cell function outcomes using the 6 year longitudinal data from the PROMISE cohort. Variables controlled for were follow-up time, waist circumference, baseline age, ethnicity, sex, ALT, physical activity, and total NEFA. X-axis values represent a percent difference in the outcome per SD increase in the fatty acid. P-values were adjusted for the BH false discovery rate, with the largest dot representing a significant (p<0.05) association.')
fig_pls <- fig('pls', 'Partial least squares (PLS) models showing the clustering of triacylglycerol fatty acids on insulin sensitivity and beta-cell function measures. The R value shown above the graph is the correlation between predicted and actual values, using cross-validation; a higher value suggests the fatty acids strongly predict the metabolic outcome. The percent explained variance of each component is shown in brackets on each axis. The solid line represents an explained variance of 100% while the dashed line represents an explained variance of 50%. Fatty acids between these lines represent variables that strongly explain the underlying structure of the data. See the Supplemental Methods for a description of PLS analysis and an explanation of interpreting this plot.')
```

@@

Recruitment for the baseline visit of the Prospective Metabolism and Islet Cell Evaluation (PROMISE) 
cohort took place between 2004-2006 from London and Toronto, 
Canada. Individuals were selected to participant if they met the eligibility 
criteria of having one or more risk factors for type 2 diabetes mellitus, 
including obesity, hypertension, family history of diabetes, and/or a history of
gestational diabetes or birth of a macrosomic infant. A total of 736 individuals
attended the baseline visit. Subsequent examinations occurred every three years,
with three examination visits completed to date (2004-2006, 2007-2009, and
2010-2013). Participants are contacted annually by telephone. The current study
sample used data on participants who did not have diabetes at baseline, who
returned for one or more of the follow-up examinations, and who had samples
available for fatty acid measurements (n=477; see the CONSORT diagram in 
`r cite_f('consort')`; a total of `r calc_n_for_visits()[1]` attended all three
visits). Metabolic characterization, anthropometric
measurements, and questionnaires on lifestyle and sociodemographics are
collected at each examination visit. Research ethics approval was obtained from
Mount Sinai Hospital and the University of Western Ontario, and all participants
provided written informed consent. Data collection methods were standardized
across the 2 centres and research nurses were centrally trained.

@@

`r cite_f('tagfa')` shows the composition of each FA in the TAG fraction (see 
`r cite_t('tagfa')` for a tabular presentation of the values). 
Three TAGFA contributed `r format_rounding(sum(tagfa$pct))`% to the total 
TAG concentration: `r tagfa$c[1]`; `r tagfa$c[2]`; and, `r tagfa$c[3]`.
`r cite_f('heatmap')` shows a heatmap of the correlation of individual TAGFA with the
outcome variables and several basic characteristics.
As expected, nearly all TAGFA had very strong positive correlations (r=
`r fa_cor$rng$TAG`) with clinically-measured TAG and moderate positive 
correlations with WC (r=`r fa_cor$rng$Waist`). There was also moderate
negative correlations with HDL (r=`r fa_cor$rng$HDL`). For the outcome variables,
the correlations for the insulin sensitivity measures were generally higher 
(HOMA2-%S: r=`r fa_cor$rng['HOMA2-%S']`, ISI: r=`r fa_cor$rng$ISI`) than 
for the beta-cell function measures (all r<0.30). A heatmap of the correlations
of individual TAGFA using mol% with the basic characteristics is shown in 
`r cite_f('heatmap_mol')` and the inter-correlation matrix of the TAGFA fraction 
is shown in `r cite_f('heatmap_tagfa')`.

@@

Both clinically-measured TAG and total TAGFA concentration had very strong
negative associations with all outcome variables. There were no significant
interactions by time, sex, or ethnicity for any of the TAGFA (data not shown).
A tabular presentation of the GEE results is shown in `r cite_t("gee_unadj")` 
for unadjusted models and `r cite_t("gee_adj")` for adjusted models.

@@

```{r fig_heatmap, fig.cap=fig_heatmap_mol}
plot_heatmap(cor_df_mol, unit = 'mol%')
```

@@

For the confounders, they were chosen based on previous literature, from
directed acyclic graph (DAG) [@Greenland1999a] recommendations, and from
QIC. The DAG recommendations were obtained from using the DAGitty software
[@Textor2011a], http://dagitty.net/. See `r cite_f('dag_is')` and 
`r cite_f('dag_bcf')`) for the DAG model. See `r cite_t('qic')` for the
comparison of various models with different covariates using QIC. While the
final GEE model selected as best fitting differed between insulin sensitivity
and beta-cell function measures, the differences in QIC values were less than 10
between many of the models, suggesting similar fit. As such, we selected the
model that had the fewest covariates and that had similar fit between the
outcome measures.

@@

>>>>

tab_qic <- stab('qic', 'Comparison of GEE model fitness for variable selection using quasi-likelihood information criteria.')
tab_basic <- tab('basic', 'Basic characteristics of PROMISE participants at each of the 3 clinic visits.')
tab_tagfa <- stab('tagfa', 'Concentration (nmol/mL) and relative percent (mol%) values of triacylglycerol fatty acids in PROMISE participants at the baseline visit (2004-2006).')
tab_gee_unadj <- stab('gee_unadj', 'Raw estimates and confidence interval values for *time*-adjusted GEE models of the association of the triacylglycerol fatty acids (mol% and nmol/mL) and total clinically-measured TAG with insulin sensitivity and beta-cell function outcomes using the 6 year longitudinal data from the PROMISE cohort. Estimates represent a percent difference in the outcome per SD increase in the fatty acid. P-values were adjusted for the BH false discovery rate, with an asterisk (*) denoting a significant (p<0.05) association.')
tab_gee_adj <- stab('gee_adj', 'Raw estimates and confidence interval values for *fully*-adjusted GEE models of the association of the triacylglycerol fatty acids (mol% and nmol/mL) and total clinically-measured TAG with insulin sensitivity and beta-cell function outcomes using the 6 year longitudinal data from the PROMISE cohort. Variables controlled for were follow-up time, waist circumference, baseline age, ethnicity, sex, ALT, physical activity, and total NEFA. Estimates represent a percent difference in the outcome per SD increase in the fatty acid. P-values were adjusted for the BH false discovery rate, with an asterisk (*) denoting a significant (p<0.05) association.')

fig_consort <- sfig('consort', 'CONSORT diagram of PROMISE participants over the 3 visits.')
fig_dag_is <- sfig('dag_is', 'Directed acyclic graphic output from the DAGitty online software for insulin sensitivity.')
fig_dag_bcf <- sfig('dag_bcf', 'Directed acyclic graphic output from the DAGitty online software for beta-cell function.')
fig_tagfa <- fig('tagfa', 'Distribution of the composition of triacylglycerol fatty acids in the baseline visit of PROMISE participants (2004-2006). Boxplots represent the median and interquartile range of the fatty acid values.')
fig_heatmap <- fig('heatmap', 'Pearson correlation heatmap of triacylglycerol fatty acids (nmol/mL) with continuous basic and metabolic characteristics of PROMISE participants from the baseline visit (2004-2006). Darker orange represents a positive correlation; darker blue represents a negative correlation.')
fig_heatmap_mol <- fig('heatmap_mol', 'Pearson correlation heatmap of triacylglycerol fatty acids (mol%) with continuous basic and metabolic characteristics of PROMISE participants from the baseline visit (2004-2006). Darker orange represents a positive correlation; darker blue represents a negative correlation.')
fig_heatmap_tagfa <- fig('heatmap_tagfa', 'Pearson correlation heatmap of the triacylglycerol fatty acids in the PROMISE participants from the baseline visit (2004-2006). The correlations of fatty acids grouped using heirarchical cluster analysis; fatty acids along the x and y axis are ordered according to this analysis. Darker orange represents a positive correlation; darker blue represents a negative correlation.')
fig_gee_unadj <- fig('gee_unadj', 'Time-adjusted GEE models of the association of the triacylglycerol fatty acids (mol% and nmol/mL) and total clinically-measured TAG with insulin sensitivity and beta-cell function outcomes using the 6 year longitudinal data from the PROMISE cohort. X-axis values represent a percent difference in the outcome per SD increase in the fatty acid. P-values were adjusted for the BH false discovery rate, with the largest dot representing a significant (p<0.05) association.')
fig_gee_adj <- fig('gee_adj', 'Fully-adjusted GEE models of the association of the triacylglycerol fatty acids (mol% and nmol/mL) and total clinically-measured TAG with insulin sensitivity and beta-cell function outcomes using the 6 year longitudinal data from the PROMISE cohort. Variables controlled for were follow-up time, waist circumference, baseline age, ethnicity, sex, ALT, physical activity, and total NEFA. X-axis values represent a percent difference in the outcome per SD increase in the fatty acid. P-values were adjusted for the BH false discovery rate, with the largest dot representing a significant (p<0.05) association.')
fig_pls <- fig('pls', 'Partial least squares (PLS) models showing the clustering of triacylglycerol fatty acids on insulin sensitivity and beta-cell function measures. The R value shown above the graph is the correlation between predicted and actual values, using cross-validation; a higher value suggests the fatty acids strongly predict the metabolic outcome. The percent explained variance of each component is shown in brackets on each axis. The solid line represents an explained variance of 100% while the dashed line represents an explained variance of 50%. Fatty acids between these lines represent variables that strongly explain the underlying structure of the data. See the Supplemental Methods for a description of PLS analysis and an explanation of interpreting this plot.')
```

@@

Recruitment for the baseline visit of the Prospective Metabolism and Islet Cell Evaluation (PROMISE) 
cohort took place between 2004-2006 from London and Toronto, 
Canada. Individuals were selected to participant if they met the eligibility 
criteria of having one or more risk factors for type 2 diabetes mellitus, 
including obesity, hypertension, family history of diabetes, and/or a history of
gestational diabetes or birth of a macrosomic infant. A total of 736 individuals
attended the baseline visit. Subsequent examinations occurred every three years,
with three examination visits completed to date (2004-2006, 2007-2009, and
2010-2013). Participants are contacted annually by telephone. The current study
sample used data on participants who did not have diabetes at baseline, who
returned for one or more of the follow-up examinations, and who had samples
available for fatty acid measurements (n=477; see the CONSORT diagram in 
`r cite_sf('consort')`; a total of `r calc_n_for_visits()[1]` attended all three
visits). Metabolic characterization, anthropometric
measurements, and questionnaires on lifestyle and sociodemographics are
collected at each examination visit. Research ethics approval was obtained from
Mount Sinai Hospital and the University of Western Ontario, and all participants
provided written informed consent. Data collection methods were standardized
across the 2 centres and research nurses were centrally trained.

@@

`r cite_f('tagfa')` shows the composition of each FA in the TAG fraction (see 
`r cite_st('tagfa')` for a tabular presentation of the values). 
Three TAGFA contributed `r format_rounding(sum(tagfa$pct))`% to the total 
TAG concentration: `r tagfa$c[1]`; `r tagfa$c[2]`; and, `r tagfa$c[3]`.
`r cite_f('heatmap')` shows a heatmap of the correlation of individual TAGFA with the
outcome variables and several basic characteristics.
As expected, nearly all TAGFA had very strong positive correlations (r=
`r fa_cor$rng$TAG`) with clinically-measured TAG and moderate positive 
correlations with WC (r=`r fa_cor$rng$Waist`). There was also moderate
negative correlations with HDL (r=`r fa_cor$rng$HDL`). For the outcome variables,
the correlations for the insulin sensitivity measures were generally higher 
(HOMA2-%S: r=`r fa_cor$rng['HOMA2-%S']`, ISI: r=`r fa_cor$rng$ISI`) than 
for the beta-cell function measures (all r<0.30). A heatmap of the correlations
of individual TAGFA using mol% with the basic characteristics is shown in 
`r cite_f('heatmap_mol')` and the inter-correlation matrix of the TAGFA fraction 
is shown in `r cite_f('heatmap_tagfa')`.

@@

Both clinically-measured TAG and total TAGFA concentration had very strong
negative associations with all outcome variables. There were no significant
interactions by time, sex, or ethnicity for any of the TAGFA (data not shown).
A tabular presentation of the GEE results is shown in `r cite_st("gee_unadj")` 
for unadjusted models and `r cite_st("gee_adj")` for adjusted models.

@@

```{r fig_heatmap_mol, fig.cap=fig_heatmap_mol}
plot_heatmap(cor_df_mol, unit = 'mol%')
```

@@

For the confounders, they were chosen based on previous literature, from
directed acyclic graph (DAG) [@Greenland1999a] recommendations, and from
QIC. The DAG recommendations were obtained from using the DAGitty software
[@Textor2011a], http://dagitty.net/. See `r cite_sf('dag_is')` and 
`r cite_sf('dag_bcf')`) for the DAG model. See `r cite_st('qic')` for the
comparison of various models with different covariates using QIC. While the
final GEE model selected as best fitting differed between insulin sensitivity
and beta-cell function measures, the differences in QIC values were less than 10
between many of the models, suggesting similar fit. As such, we selected the
model that had the fewest covariates and that had similar fit between the
outcome measures.

@@

## Supplemental Tables

## Supplemental Figures
